---
title: Functions
author: "Brian A. Fannin"
output:
  revealjs::revealjs_presentation:
    center: no
    css: ./css/revealOpts.css
    reveal_plugins:
    - notes
    - zoom
    self_contained: no
    theme: solarized
    transition: slide
---

```{r echo=FALSE}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(knitr))
knitr::opts_chunk$set(
    warning=TRUE
  , error=TRUE
  , echo=TRUE
  , message=FALSE
  , fig.height = 4.5
  , fig.pos="t"
  , collapse = TRUE
)
knitr::opts_knit$set(root.dir = normalizePath('../'))
```

# Yeah, functions

## Why is this amazing?

```{r eval=FALSE}
some_var <- sqrt(pi) + 5
some_function <- function(x) {
  sqrt(x) + 5
}
some_function(pi)
```

## Here's what's amazing

1. Functions are __objects__
2. Functions can return any value
3. 1 and 2 mean that functions can return other functions
4. Functions are (usually) __vectorized__
4. Functions may be altered after they're created

# What is a function made of?

## A simple function

```{r}
layer <- function(loss, attachment, limit, share){
  pmin(limit, pmax(0, loss - attachment))
}
```

## The elements of a function

1. arguments
2. body
3. environment

# Parameters

## Parameters

```{r}
formals(layer)
```

Note that it _doesn't_ tell you what data types they are. R is not strongly typed.

## Matching arguments

1. Fully named arguments
2. Partial matching
3. Position

## Matching arguments

```{r}
myFunc <- function(mojo, monkey, banana){
  print(match.call())
}

myFunc(mojo = 1, mo = 2, 3)
myFunc(mo = 2, mojo = 1, 3)
myFunc(mo = 1, mo = 2, 3)
myFunc(1, 2, 3)
```

## Default values

```{r}
myFunc <- function(a, b = 15){
  a + b
}
myFunc(3)
```

I didn't supply a value for the `b` argument, so we just used the default.

## Check inputs

You can use the `missing()` function for this.

```{r}
myFunc <- function(mojo, monkey, banana){
  if (missing(banana)) {
    warning("I can't think without bananas")
    banana = monkey + 5
  }
}
myFunc(1, 2, 3)
myFunc(1, 2)
```

## Aside

* `stop()`: This is an error. The program can't go on.
* `warning()` - This tells the user that something happened and they might not like it, but it's not a critical issue.
* `message()` - Just sayin' something you might want to hear.

## Missing and defaults

```{r}
myFunc <- function(mojo, monkey, banana = 3){
  if (missing(banana)) {
    warning("I can't think without bananas")
    print(banana)
  }
}
myFunc(1, 2, 3)
myFunc(1, 2)
```

## Restrict inputs to set choice

```{r}
myFunc <- function(x, dist = c("lognormal", "gamma")){
  dist = match.arg(dist)
  print(dist)
}
myFunc(1:10, "lo")
myFunc(1:10, "jo")

myOtherFunc <- function(x, dist = c("lognormal", "gamma")){
  print(dist)
}
myOtherFunc(1:10, "lo")
myOtherFunc(1:10, "jo")
```

## Defaults and missing arguments

* If you _must_ know what the user intends, _or_ if the default requires some logic, don't add a default.
* If you're fine with common default parameters (i.e. the base of the log function is e or 10), add a default.
* Don't add a default and check for a missing value. It's confusing and unnecessary.

## Triple dot

`...` is used to pass unspecified arguments to other functions. Typically used if your function sits between the user and some other function like `plot`

```{r, echo=TRUE}
dfData <- data.frame(PolicyYear = 2001:2010, Premium = rnorm(10))
myPlotFunc <- function(dfData, ...){
  plot(dfData$PolicyYear, dfData$Premium, ...)
}
```
```{r eval=FALSE}
myPlotFunc(dfData, pch=18)
myPlotFunc(dfData, pch=23)
```

```{r}
formals(myPlotFunc)
```

# Return value

## What does a function return?

Anything.

## return()

* `return()` isn't required. 
* A function will always return the last evaluated expression.
* `return()` is often used for _abnormal_ returns

```{r}
myFunc <- function(x){
  if (is.na(x)) return("Value missing")
  
  sum(x)
}
```

## Create a list of functions.

```{r}
lstMeans <- list(arithmetic = mean, median = median)
nums <- 1:10
lstMeans[['arithmetic']](nums)
which_mean <- 'median'
lstMeans[[which_mean]](nums)
lapply(lstMeans, function(x) x(nums))
```

## What's a closure?

A closure is a function which contains data.

```{r}
my_rnorm <- function(mean, sd){
  function(N){
    rnorm(N, mean, sd)
  }
}
an_rnorm <- my_rnorm(mean = 10, sd = 3)
```

```{r }
set.seed(1234)
my_norms <- an_rnorm(100)
set.seed(1234)
base_norms <- rnorm(100, 10, 3)
identical(my_norms, base_norms)
```

## Where did the data go?

It's in the enclosing environment of the function.

```{r}
an_rnorm <- my_rnorm(mean = 10, sd = 3)
another_rnorm <- my_rnorm(mean = -30, sd = 100)
ls(environment(an_rnorm))

func_data <- ls(environment(an_rnorm))
lapply(func_data, get, environment(an_rnorm))

as.list(environment(an_rnorm))
as.list(environment(another_rnorm))
```

## Exercise

* $\mu=log(E[X])-\sigma^2$
* $\sigma^2=ln(1+cv^2)$

Create a function called LognormalParams, which will translate the mean and either the CV or standard deviation or variance into parameters for the lognormal distribution. Test it using the following statements:

```{r eval=FALSE}
LognormalParams(10e3, 3.0)
LognormalParams(c(10e3, 20e3), 3.0)
LognormalParams(c(10e3, 20e3), "three")
LognormalParams(10e3)
LognormalParams(CV = 4, 10e3)
```

## Another exercise

## Solution

```{r}
LognormalParams <- function(mean, CV){
  
}
```

# If we have time

## Replacement functions

(These examples are virtually identical to the ones in Hadley's "Advanced R")

```{r}
`mojo<-` <- function(x, value){
  x <- x + value
} 
gonzo <- 1:10
gonzo
mojo(gonzo) <- 5
gonzo
```

## Replacement functions (2)

```{r}
`mojo<-` <- function(x, position, value){
  x[position] <- x[position] + value
  x
} 
gonzo <- rep(1, 10)
gonzo
mojo(gonzo, 2) <- 5
gonzo
mojo(gonzo, 2:3) <- 5
gonzo
```

## Infix functions

```{r}
`%l^%` <- function(a, b){
  log(sqrt(a +b))
}
5 %l^% 3
log(sqrt(5+3))
```

# NSE

## NSE

I really don't want to talk about this

##

1. `parse()` will convert text into an expression.
2. `eval()` will evaluate an expression. 
3. `deparse()` will convert an expression back into text (but not really).

##

```{r }
my_str <- "1 + 1"
my_expr <- parse(text = my_str)
my_expr
class(my_expr)
```

## Deparse is crazy

```{r }
my_other_str <- deparse(my_expr, width = 500)
my_other_str

my_other_str <- as.character(my_expr)
my_other_str
as.character(attr(my_expr, "wholeSrcref"))
my_other_str
```

## eval

```{r }
eval(parse(text = my_expr))
my_expr <- "1 + 2"
eval(parse(text = my_expr))
```
