---
title: Introduction
author: "Brian A. Fannin"
output:
  revealjs::revealjs_presentation:
    center: no
    css: ./css/revealOpts.css
    reveal_plugins:
    - notes
    - zoom
    self_contained: no
    theme: solarized
    transition: slide
---

```{r echo=FALSE}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(knitr))
knitr::opts_chunk$set(
    warning=FALSE
  , error=FALSE
  , echo=FALSE
  , message=FALSE
  , fig.height = 4.5
  , fig.pos="t"
  , collapse = TRUE
)
knitr::opts_knit$set(root.dir = normalizePath('../'))
```

# Hello!

## Answer one of these questions

* What's your favorite pizza topping?
* Who is your favorite 19th century president, other than Lincoln?
* Where is your favorite place to take a vacation?

## Our day: software not statistics

By the end of the day you will

* Write robust functions, capable of gracefully handling unexpected inputs
* Create a package, which could be submitted to CRAN
* Create tests to ensure that your code is much less likely to break
* Understand how to use git to develop safely, collaboratively and with discipline
* Use R and markdown to create documents which express complex statistical ideas in a way that's easy to understand
* Create a web-based application which supports a rich, interactive user experience

## Agenda

```{r results='asis'}
tblAgenda = tribble(
  ~Start, ~End, ~What, ~`How Long`
  , "9:00", "9:15", "Introductions", "15 min"
  , "9:15", "10:15", "More fun with functions", "1 hour"
  , "10:15", "10:30", "Packages 1 - getting started", "15 min"
  , "10:30", "11:00", "AM break", "30 min"
  , "11:00", "12:30", "Packages 2 - here we go", "90 min"
  , "12:30", "1:30", "Lunch", "1 hour"
  , "1:30", "2:15", "Get started with git", "45 min"
  , "2:15", "3:00", "Markdown", "45 min"
  , "3:00", "3:30", "PM break", "30 min"
  , "3:30", "4:15", "Shiny", "45 min"
  , "4:15", "5:00", "Padding", "45 min"
)
tblAgenda %>% kable()
```

