---
title: "Packages - Part 2"
author: "Brian A. Fannin"
output:
  revealjs::revealjs_presentation:
    center: no
    css: ./css/revealOpts.css
    reveal_plugins:
    - notes
    - zoom
    self_contained: no
    theme: solarized
    transition: slide
---

```{r echo=FALSE}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(knitr))
knitr::opts_chunk$set(
    warning=FALSE
  , error=FALSE
  , echo=FALSE
  , message=FALSE
  , fig.height = 4.5
  , fig.pos="t"
  , collapse = TRUE
)
knitr::opts_knit$set(root.dir = normalizePath('../'))
```

# Workflow

## Workflow

1. Create a new project in RStudio
2. Edit DESCRIPTION file
4. Set build parameters
2. Enter a function
3. Document with roxygen
4. Add data?
5. Create tests
6. Build
7. Repeat steps 2-8

## Quick tour of what RStudio created

## License

Your package needs one, especially if the software is proprietary.

## Dependencies

One of two places

1. DESCRIPTION file
2. NAMESPACE

## Dependencies in DESCRIPTION

* depends - The package will be placed on the search path
* imports - The package will be available, but not on the search path
* suggests - The package does not need this to run, but some functionality may be enhanced
* improves - This is rare. This means that the package enhances functionality in another package

Use depends sparingly.

## NAMESPACE

We'll get to that in a bit

## Exercise

Edit the DESCRIPTION file to add a suggestion for `ggplot2` (and possibly MASS) and build the package. If you haven't changed any of the defaults, the build should take about 3 seconds and produce no errors.

# Functions

## Writing functions

This is the easy bit. Often, you're just taking code from a script and dumping it into a package. It needs clean up, but writing functions isn't super hard.

## Exercise

Create a function in the package which will translate mean and CV into lognormal parameters.

While you're at it, create a function which will translate mean and CV into gamma parameters.

$\alpha=1/CV^2$

$\beta=1 / (mean * CV^2)$

# Documentation

## roxygen

Documentation is the tricky - but super essential! - part. Happily, roxygen will take care of most of this.

## roxygen

```{r}
tbl_roxygen <- tribble(
  ~Directive, ~`What It Does`
  , "@title", "Typically the name of your function"
  , "@description", "Short description of the function's purpose"
  , '@param', "One entry for each argument to the function"
  , '@details', "Longer description of the function"
  , '@return', "What value does it return?"
  , '@importFrom', "Does the function need functions from another package?"
  , '@include', "Does this function depend on other files in your package?"
  , '@export', "Should this function be exported?"
  , '@examples', "Examples of how to use your function"

)
tbl_roxygen %>% kable()
```

Poorly documented code will not be accepted on CRAN!

## Namespace and collation

@importFrom will add entries to the NAMESPACE file.
@include will ensure that the collation sequence is accurate.

What's collation? When building the package, things happen sequentially. In your package, if a function in in file A depends on another function in a file B, file B must be processed first. Otherwise, it won't be understood when the package compiler processes file A.

## Exercise

Document your functions and rebuild your package. Gaze in awe at the entries in the help file.

# Testing

## Testing is much, much more important than the code you write

> A bug is a test that you failed to write

##

```{r eval=FALSE, echo=TRUE}
devtools::use_testthat()
devtools::use_test("my test")
```

## Great expectations

```{r eval=FALSE}
context("lognormal")

test_that("Check inputs", {
  expect_error(LognormalParams())
  expect_error(LognormalParams(), "No mean was specified")
  expect_error(LognormalParams(3), "No CV was specified")

  # and so on
  
  expect_silent(LognormalParams(5, 4))

})
```

## Exercise

Write a test suite for your function(s). Be sure to give your functions inputs that you __know__ will cause an error.

## One more item

To be sure that we're cool, we're going to add one more function.

Add a function which will accept a vector of sample data and value(s) for a range around the mean and/or CV and whatever parameters you think are appropriate. The function will return data which indicates the likelihood function at combinations of mean and CV.

For example:

```{r eval=FALSE}
my_likelihood <- CalcLikelihood(somedata, mean_range = c(50e3, 150e3), CV_range = c(0.3, 1.8))
```

# If we have time

## Data in a package

## Why data?

* Data is useful to illustrate functionality
    * ChainLadder has several triangles
* Data may be the reason for the package
* Data could be used for testing

## Three ways to get data into a package

1. In the data folder
2. In the system files
3. Within the package environment

## In the data folder

```
devtools::use_data_raw()
```

## One more exercise

Add some data, which  may be used to illustrate the likelihood function.
