---
title: "Shiny"
author: "Brian A. Fannin"
output:
  revealjs::revealjs_presentation:
    center: no
    css: ./css/revealOpts.css
    reveal_plugins:
    - notes
    - zoom
    self_contained: no
    theme: solarized
    transition: slide
---

```{r echo=FALSE}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(knitr))
knitr::opts_chunk$set(
    warning=FALSE
  , error=FALSE
  , echo=FALSE
  , message=FALSE
  , fig.height = 4.5
  , fig.pos="t"
  , collapse = TRUE
)
knitr::opts_knit$set(root.dir = normalizePath('../'))
```

# What

## What is shiny?

* An R interface to java script libraries
* Allows for user interfaces
* Support complex interactivity

## Basic structure

Two parts:

1. ui - contains user interface elements
2. server - how does the application respond to changes in the ui?

It's also possible to define an app using the `shinyApp()` function, but we won't do that here.

## Let's begin

Create a new shiny project in RStudio.

# ui

## Just a collection of ui items and some structure

* Select input
* Date input
* Text input
* Numeric sliders
* Radio buttons

## Layout and themes

This can get crazy. Spend a bit of time learning about Bootstrap and responsive web design. For now, just focus on elements.

## Exercise

* Add a radio button to select between the gamma and lognormal distributions
* Add a slider for ranges of CV and Mean values

# server

## 3 basic elements

1. A render function
2. A reactive value
3. An observer

## Exercise

* Add some plot output to the server using the renderPlot function.

## Reactivity

Enter the following line before starting your app:

```{r}
options(shiny.reactlog=TRUE) 
```

Press Ctrl+F3 when your app starts running. Cool!

## Reactivity vs. Observables

* A reactive expression is meant to - wait for it - react to its inputs having changed. When that happens, it will recalculate and return a value
* An observable is meant to _take action_ when it notices something happening. It does _not_ return a value.

A reactive is a __noun__
An observable is a __verb__

## Validation

```{r eval=FALSE}
my_reactive <- reactive({
  validate(
        need(logged_in(), "You're not logged in")
      , need(length(df_my_data()) != 0, "No data was found.")
    )
  # Now do something
})
```

## Exercise

* Add validation to ensure that the parameters exist before you try to plot.
